01_Vegetation_cover

### 2022-09-08 01_Species_abundance.Rmd
add swiss data

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(readxl)
library(Taxonstand)
```

## To do

@@

```{r}
# load data
spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)

# Make tidy
spec_a <- spec_a_raw %>% 
  pivot_longer(3:19, names_to = "plot", values_to = "abun") %>% 
  rename(sitename = Site, species = Species) %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  # create direction and distance column
  mutate(direction = case_when(str_detect(plot, "N") ~ "North",
                               str_detect(plot, "S") ~ "South",
                               str_detect(plot, "E") ~ "East",
                               str_detect(plot, "W") ~ "West",
                               str_detect(plot, "NE") ~ "Northeast",
                               str_detect(plot, "NW") ~ "Northwest",
                               str_detect(plot, "SW") ~ "Southwest",
                               str_detect(plot, "SE") ~ "Southeast",
                               plot == "Center" ~ "Center"),
         distance = case_when(plot %in% c("N...4", "E...5", "S...6","W...7") ~ 
                                "1.5-3 meter",
                              plot %in% c("N...8", "E...9", "S...10", "W...11")  ~ 
                                "3-6 meter",
                              plot %in% c("N...12","NE","SE","NW","SW", "E...14", "S...16", "W...18") ~ 
                                "6-10 meter",
                              plot == "Center" ~ "0 meter")) %>% 
  # create unique id for plot in site
  group_by(sitename, direction, distance) %>% 
  mutate(plot = cur_group_id())

@@

>>>>

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(readxl)
library(LCVP)
library(lcvplants)
```

# To do

# Scotland

@@

```{r}
# load data
spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)
canop_a_raw <- read_xlsx("Data/Canopy_zone_A.xlsx")

canop_a <- canop_a_raw %>% 
  pivot_longer(-sitename, names_to = "species", values_to = "abun") %>% 
  mutate(direction = "Canopy", distance = "0-10 meter")
# Make tidy
spec_a <- spec_a_raw %>% 
  pivot_longer(3:19, names_to = "plot", values_to = "abun") %>% 
  rename(sitename = Site, species = Species) %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  # create direction and distance column
  mutate(direction = case_when(str_detect(plot, "N") ~ "North",
                               str_detect(plot, "S") ~ "South",
                               str_detect(plot, "E") ~ "East",
                               str_detect(plot, "W") ~ "West",
                               str_detect(plot, "NE") ~ "Northeast",
                               str_detect(plot, "NW") ~ "Northwest",
                               str_detect(plot, "SW") ~ "Southwest",
                               str_detect(plot, "SE") ~ "Southeast",
                               plot == "Center" ~ "Center"),
         distance = case_when(plot %in% c("N...4", "E...5", "S...6","W...7") ~ 
                                "1.5-3 meter",
                              plot %in% c("N...8", "E...9", "S...10", "W...11")  ~ 
                                "3-6 meter",
                              plot %in% c("N...12","NE","SE","NW","SW", "E...14", "S...16", "W...18") ~ 
                                "6-10 meter",
                              plot == "Center" ~ "0 meter")) %>% 
  bind_rows(canop_a) %>% 
  # create unique id for plot in site
  group_by(sitename, direction, distance) %>% 
  mutate(plot = cur_group_id())

@@

# Switzerland

Land cover data was downloaded from:
<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability.html>

To decrease file size, I only selected relevant columns: the one
containing coordinates and land cover data using the following code:

```{r eval = FALSE}
# ag_b_00_03_37_area_csv <- read_delim(
#     "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data analysis/PhD/Validation-study/Data/Swiss_LC/ag-b-00.03-37-area-csv.csv",+delim = ";",
#     escape_double = FALSE,
#     trim_ws = TRUE
#   )
# 
# LC <- ag_b_00_03_37_area_csv %>%
#   dplyr::select(RELI, E_COORD = E, N_COORD = N, LC18_6, LC18_27)
# 
# write.csv(LC, file = "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data
#           analysis/PhD/Validation-study/Data/Swiss_LC/Swiss_LC18.csv")

In this file I found the CRS:
<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability/swiss-land-use-statistics.assetdetail.20104775.html>\
"Les coordonnées « E_COORD » et « N_COORD » des hectares dans le système
de coordonnées suisse MN95 (7 chiffres). Les coordonnées représentent le
point d'échantillonnage correspondant (intersections des coordonnées
hectométriques)."

Import Swiss land cover (100m resolution) Swiss_LC18.csv

-   Set CRS code to EPSG:4150 Rasterize this layer using the GDAL
    rasterize (vector to raster) tool
-   Set burnin to LC18_27
-   Horizontal and vertical resolution to 100 m
-   export and set resolution to 25 m for merge with forest cover data

Import forest cover data in ArcGIS pro reproject to EPSG:4150 and export
raster(folder called reproj)/ Import to QGIS set layer crs to EPSG:4150

Import coordinates of pollen plots (Pollen_site_coord.csv).

Use Zonal histogram to compute pixels in zone attributed to certain land
cover type Export as Veg_cover_zoneC.csv and Veg_cover_zoneB.csv

Import coordinate of all BDM sites Use Sample raster values to compute
landcover types for vegetation plots. Export as Veg_type_plots_csv

```{r}
zone_b_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneB_swiss.csv")
zone_c_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneC_swiss.csv")
LC_labels <- read_xlsx("Data/Swiss_LC/be-b-00.03-37-area-var-v01.xlsx",
                       sheet = "Codes LC_27")
FC_labels <- read.csv("Data/Swiss_LC/Forest_cover_labels.csv") %>% 
  rename(CODE = category)
LC_labels <- LC_labels %>% 
  dplyr::select(CODE, LABEL_DE) %>% 
  bind_rows(FC_labels)

zone_b <- zone_b_raw %>% 
  # make tidy
  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
  filter(!value == 0) %>% 
  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
  # filter land cover type that are not vegation
  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
                           "Versteinte Flächen","Befestigte Flächen",
                           "Gemischte Kleinstrukturen")) %>%
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov = value/sum(value)) %>%
  group_by(sitename, LABEL_DE) %>%
  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))

zone_c <- zone_c_raw %>% 
  # make tidy
  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
  filter(!value == 0) %>% 
  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
  # filter land cover type that are not vegation
  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
                           "Versteinte Flächen","Befestigte Flächen",
                           "Gemischte Kleinstrukturen")) %>%
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov = value/sum(value)) %>%
  group_by(sitename, LABEL_DE) %>%
  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))

# merge zone b and c
veg_cover <- zone_b %>% 
  full_join(zone_c, by = c("sitename", "LABEL_DE"), 
            suffix = c("_b","_c")) %>% 
  replace_na(list(veg_cov_b = 0))
```

```{r join veg type data and species data}
# Veg type of BDM sites
veg_type <- read.csv("Data/Swiss_LC/FC_LC_type_swiss.csv")

# Join with labels
veg_type <- veg_type %>% 
  left_join(LC_labels, by = c("SAMPLE_1" = "CODE"))

# Load in species data
bdm_sp <- readRDS("RDS_files/01_Species_abundance_Swiss.rds")

bdm_sp <- bdm_sp %>% 
  # filter stuff that is not a plant
  filter(!stand.spec == "No plants") %>% 
  # filter unindentified
  filter(!stand.spec == "Unindentified") %>% 
  # join with vegetation type data
  left_join(veg_type, by = c("sitename" = "aID_STAO")) 
# create table with genus and family to join after summarizing
genus_fam <- bdm_sp %>% 
  ungroup() %>% 
  dplyr::select(family, genus, stand.spec) %>% 
  distinct
```

```{r}
spec_cover <- bdm_sp %>% 
  # calculate average species cover per vegetation type using the plots belonging to that vegetation type
  group_by(sitename, LABEL_DE, stand.spec) %>% 
  summarise(abun = mean(abun)) %>% 
  # join with vegetation cover data
  right_join(veg_cover, by = c("sitename", "LABEL_DE")) %>% 
  # weight species abundance by veg cover
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c)

# print table with missing species data for vegetation types
missing_vegtype <- spec_cover %>% 
  filter(is.na(stand.spec)) %>% 
  group_by(sitename, LABEL_DE) %>% 
  summarise(cov_b = round(sum(veg_cov_b), 2), 
            cov_c = round(sum(veg_cov_c), 2)) %>%
  filter(!cov_b == 0|!cov_c == 0)

# for missing data, use species data from other plots
missing <- bdm_sp %>% 
  filter(LABEL_DE %in% missing_vegtype$LABEL_DE) %>% 
  group_by(LABEL_DE, stand.spec) %>% 
  summarise(abun = mean(abun))
spec_cover_na <- spec_cover %>% 
  filter(is.na(stand.spec)) %>%
  dplyr::select(sitename, LABEL_DE, veg_cov_b, veg_cov_c) %>% 
  left_join(missing, by = "LABEL_DE") %>% 
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) 
spec_cover_bc <-  spec_cover %>% 
  filter(!is.na(stand.spec)) %>%
  bind_rows(., spec_cover_na) %>% 
  arrange(sitename, LABEL_DE, stand.spec) %>% 
  dplyr::select(sitename, stand.spec, spec_abun_b, spec_abun_c) %>% 
  ungroup() %>% 
  left_join(genus_fam, by = "stand.spec")

saveRDS(spec_cover_bc, "RDS_files/01_Species_abundance_zoneBC_swiss.rds")
```
